#Liver Disease Prediction 

###Liver Diseases have been a widespread issue in many countries. Third World countries suffer from a sparsity of doctors, leading to difficulty in diagnosing patients. With the advent of machine learning technology, we can now try to automate the process of diagnosing patients based on their test results. This saves the doctors and patients a lot of time, and those that are in need of attention can be tended to. This is a simple attempt to solve this calssification problem using different algorithms.
The data set 
Calling the required libraries:

```{r, results='hide'}
library(ggplot2) 
library(readr)
library(psych)
library(corrplot)
library(ggpubr)
library(caret)
```

Reading the dataset file:
```{r, results='hide', include=FALSE}
liver_data <- read_csv("indian_liver_patient.csv")
```

Now we explore the dataset to find anomalies and get a better understanding of what course of action to follow:

```{r}
str(liver_data) 
summary(liver_data) 
```

From these operations we see that there are 10 attributes (dataset being our target variable), and 583 observations. We also see that there are four null values, which need to be dealt with before model building. Hence eliminating the null values:

```{r}
liver_data <- liver_data[complete.cases(liver_data), ]
#Change 2 to 0 in target variable 
liver_data[liver_data$Dataset==2,]$Dataset<-0
```

Verifying whether all the anamolies are dealt with:
```{r}
str(liver_data)
summary(liver_data)
```

The data is now ready. Exploring further to better understand the dataset.
```{r}
#Age
par(mfrow=c(1,2))
hist(liver_data$Age, main = 'Age distribution')
boxplot(liver_data$Age, main="Boxplot of Age")

#Gender
par(mfrow=c(1,1))
barplot(table(liver_data$Gender))

#Dataset(target variable)
par(mfrow=c(1,2))
barplot(table(liver_data$Dataset))
par(mfrow=c(1,1))

#Evaluating the number of healthy and unhealthy patients in the dataset(1=unhealthey, 0=healthy)
table(liver_data$Dataset)
```

From these plots we can see that there are more males than females in the data set, and the most of the people are between 40-50 years old.

Evaluating the correlations between the features

```{r}
# Non-numeric fields are not considered
cor_matrix <- cor(liver_data[,-c(2,11)]) 
cor_matrix
corrplot(cor_matrix,title = "Correlation Matrix",tl.col = "black",type = "upper")
```

Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients.On the right side of the correlogram, the legend color shows the correlation coefficients and the corresponding colors. From the plot it is eveident that there is a high correlation between the following pair : Total_Bilrubin, Direct Bilrubin; Alamine_Amniotransferase, Aspartate_Amniotranferase; Total_Proteins, Albumin. Exploring these correlations further.
```{r}
par(mfrow=c(2,2))
ggscatter(data = liver_data, x = "Total_Bilirubin", y = "Direct_Bilirubin", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Total Bilirudbin", ylab = "Direct Bilirubin")

ggscatter(data = liver_data, x = "Total_Protiens", y = "Albumin", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Total Protiens", ylab = "Albumin")

ggscatter(data = liver_data, x = "Alamine_Aminotransferase", y = "Aspartate_Aminotransferase", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Alamine_Aminotransferase", ylab = "Aspartate_Aminotransferase")

ggscatter(data = liver_data, x = "Albumin", y = "Albumin_and_Globulin_Ratio", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          
          xlab = "Albumin", ylab = "Albumin_and_Globulin_Ratio")
```
These pairs are highly correlated and thus only 3 out of the 6 features need to be selected as they would have the same impact on the model. Eliminating Gender, Direct_Bilrubin, Alkaline_Phosphotase, Asparate_Amino_Transferase, in an attempt to build more robust models.
```{r}
#Select the columns we need as predictors
str(liver_data)
liver_data <- liver_data[,c(1,3,6,8,9,10,11)]
colnames(liver_data)
```

Now dividing the dataset into training and testing sets.
```{r}
# 70% of the sample size
smp_size <- floor(0.70 * nrow(liver_data))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(liver_data)), size = smp_size)
train_data <- liver_data[train_ind, ]
test_data <- liver_data[-train_ind, ]

str(train_data)
str(test_data)
str(liver_data)
```
Now that the data has been prepared, the model construction process can be initiated. Since this is a classificaiton problem, the implementation of KNN, CART, SVM, and Logistic Regression.



